---
title: "covid 19"
author: "Jordan Rose"
date: "10/19/2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 
### This is contained in a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Question 1 - Is dying independent of sex




Table of patients:
```{r}

# cross tabulation, patient state by gender

with(kRdata, table(Sex, result))

```

## We create two bar charts - one based on frequency and the other on percent

```{r}
barchartGC(~result + Sex,data=kRdata, main ="Frequency Bar Chart")
```

  
### this bar chart is displaying in total frequency terms our data.  Clearly there are a lot less dead people than alive people as is reflected by the sizes of the bars in the display.  
```{r}
barchartGC(~result + Sex,data=kRdata, type = "percent", main ="Frequency Bar Chart")
```

Based off the this chart it shows that 100% of the alive group, about 59% are women and 43% are men. Of the dead group, about 40% are women and about 62% are men. 


## Numerical results


###Numerical Summaries  

```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```

Of the people alive 56% are female and 46% are male and for the people that are dead, 36% of them are female and 64% of them are men. 

## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```
```{r}
(1/4)/(3/9)
```

## Conclusions

There is an increased chance of men dieing from covid19 than there is of women dieing. This was concluded through the use of the fisher exact thest where the same estimates equal the odd ratio equal 2.3                                                                      